MYC overrides HIF to regulate proliferating primary cell metabolism in hypoxia

### 2022-07-19 manuscript/manuscript.Rmd
work on stable isotope tracing results and figure

## Isotope tracing demonstrates less incorporation of glucose into downstream pathways

To investigate intracellular metabolic reprogramming in hypoxic cells, we performed ^13^C stable isotope tracing with [U-^13^C~6~]-glucose, [1,2-^13^C~2~]-glucose, and [U-^13^C~5~]-glutamine. Isotopic enrichment of downstream metabolites in glycolysis and the TCA cycle were determined by LC-MS.

<!-- discuss LF v. PASMC -->

>>>>

## Isotope tracing reveals altered substrate utilization in hypoxia

To investigate intracellular metabolic reprogramming in hypoxic cells, we performed ^13^C stable isotope tracing with [U-^13^C~6~]-glucose, [1,2-^13^C~2~]-glucose, and [U-^13^C~5~]-glutamine. Isotopic enrichment of downstream metabolites in glycolysis and the TCA cycle were determined by LC-MS (**Figs S3, S4**). Overall, relatively small changes in the patterns of isotope incorporation were observed following hypoxia or BAY treatment. The most substantial differences were observed in pyruvate (PYR), the terminal product of glycolysis, and citrate (CIT), a central metabolic node in TCA and fatty acid metabolism (**Figs \@ref(fig:m3)A-C**). Both hypoxia and BAY treatments decreased incorporation of glucose-derived carbon into pyruvate (**Fig \@ref(fig:m3)A**) (*i.e.,* the unlabeled, or M0, fraction was greater). This suggests a greater contribution from an unlabeled carbon source, such as extracellular pyruvate, lactate, or alanine, than from glucose to the intracellular pyruvate pool following PHD inhibition.

Total citrate labeling from [U-^13^C~6~]-glucose was unchanged across the treatment conditions (**Fig \@ref(fig:m3)B**). As expected, we observed decreased M2 and M4 citrate isotopes, consistent with decreased pyruvate dehydrogenase activity in hypoxia. Interestingly, we observed increased M3 and M5 citrate isotopes. Pyruvate carboxylase catalyzes the carboxylation of pyruvate to oxaloacetate after which all three pyruvate carbons are incorporated into citrate by citrate synthase. Thus, this labeling pattern suggests a more prominent contribution of pyruvate carboxylase to sustain TCA cycle anaplerosis despite pyruvate dehydrogenase inhibition following HIF-1α activation. By contrast to glucose labeling, much less citrate was labeled by glutamine with hypoxia or BAY with a more pronounced effect of BAY treatment (**Fig \@ref(fig:m3)C**), suggesting a less important contribution of glutamine to TCA anaplerosis under these conditions. In addition, the overall fraction of M5 citrate resulting from reductive carboxylation of glutamine-derived α-ketoglutarate was low (< 7%) (**Fig \@ref(fig:m3)D**). Although a hypoxia-mediated increase in M5 citrate was observed, the overall fraction was much less than the 10-20% levels previously reported in cancer cells [@wise2011; @metallo2011].

The stable isotope labeling patterns in PASMCs were generally similar to LFs (**Figure S5**). The most notable differences between LF and PASMC labeling were observed in citrate. Compared with LFs, a much lower fraction of total citrate was labeled by glucose in PASMCs. Less activity of pyruvate carboxylase in these cells was suggested by decreased M3 and M5 citrate isotopes after glucose labeling. Interestingly, the M5 citrate fraction in PASMCs was more consistent with previous reports from the cancer literature (**Fig \@ref(fig:m3)D**), consistent with a more important role for glutamine metabolism for biomass synthesis in these cells.

## Glycolytic flux in hypoxia is closely coupled to cell growth rate
